The Name of the Title is Hope

### 2020-03-20 conclusion.tex
grammar fixes

We have shown an experimental setup to suss out when input-output examples help
users in a type-directed synthesis environment.
%
Overall our results did not show a statistical significance ($p = 0.14$);
however, we did find some interesting data points for further research.
%
Specifically, one task did conclude a statistically significant result ($p =
0.03$), Task 2.
%
We offered a theory explaining it and future experiments that could falsify our
claim.
%
We found that users did rely on an interpreter more in the absence of examples
to offer a "feel" for the candidate programs.

>>>>

We have shown an experimental setup to suss out when input-output examples help
users in a type-directed synthesis environment.
%
Overall our results did not show a statistical significance ($p = 0.14$);
however, we did find some interesting data points for further research.
%
Specifically, one task did conclude a statistically significant result ($p =
0.03$), Task~2.
%
We offered a theory explaining it and future experiments that could falsify our
claim.
%
We found that users did rely on an interpreter more in the absence of examples
to offer a "feel" for the candidate programs and to lookup documentation.
%
We believe our study was sufficiently underpowered to be unable to make a
definitive claim on the overall utility of examples in a type-directed synthesis
environment.

### 2020-03-20 discussion.tex
grammar fixes

\subsection{Hope for examples}
Although the study failed to show that examples aided candidate selection in
the general case, it's possible that in specific case it makes a large difference.
%
A Wilcoxon Rank-Sum test on the data from Task 2 shows that the \examples
group was faster with a confidence of $p=0.03$.
%
While this could just be noise, the correct solution for Task 2 was 10th in
the list while Tasks 1 and 2 had their solution in the first two spots.
%
Task 3 had no solution.
%
If examples speed up candidate assessment, we'd expect this effect to be
magnified by the number of candidates assessed.
%
Because users had to sift through 9 incorrect solutions to get to the correct
one, it makes sense that Task 2 showed the most dramatic effect.
%
A study that randomizes solution location rather than preserving the synthesizer's ordering
could test this hypothesis.

>>>>

\subsection{Hope for examples}
Although the study failed to show that examples aided candidate selection in
the general case, it's possible that in specific cases examples makes a large difference.
%
A Wilcoxon Rank-Sum test on the data from Task 2 shows that the \examples
group was faster with a confidence of $p=0.03$.
%
While this could just be noise, the correct solution for Task 2 was 10th in
the list while Tasks 1 and 2 had their solution in the first two spots.
%
Task 3 had no solution.
%
If examples speed up candidate assessment, we'd expect this effect to be
magnified by the number of candidates assessed.
%
Because users had to sift through 9 incorrect solutions to get to the correct
one, it makes sense that Task 2 showed the most dramatic effect.
%
A study that randomizes solution location rather than preserving the synthesizer's ordering
could test this hypothesis.

### 2020-03-20 intro.tex
grammar fixes

Imagine asking your computer to write a function for you and it works on the
first try.
%
Well, program synthesis isn't quite there.
%
But getting there will necessarily involve some back and forth between the user
and computer.
%
Often the function specification is less precise the than function requested.
%
In this case, a program synthesizer can come up with any number of solutions
that all meet the specification, to varying degrees of utility.

Our work is based on \hoogleplus, a program synthesis engine for the
Haskell programming language \cite{hoogle_plus_2020}.
%
A user inputs a type signature of a function, like \texttt{a -> [Maybe a] -> a}
(read: ``A function that takes a thing and a list of optional-things, and
returns a thing'').
%
\hoogleplus then looks for Haskell programs that use library calls (from
a small, builtin set of functions) that satisfies the type signature.
%
Note that there may be an infinite number of so-called \emph{inhabitants} of a type.
%
This presents a search problem for users of the system.

>>>>

Program Synthesis works by taking a specification and generating a program that
implements the specification.
%
In a sense it is the holy grail of computer science.
%
There is a plethora of work trying to go straight to the goal.
%
Some use input-output examples as their specification
\cite{Feser_Chaudhuri_Dillig_2015}.
%
Others use types and examples \cite{Osera_Zdancewic_2015}.
%
Still others use just types \cite{hoogle_plus_2020}.

However, in lieu of waiting for some graduate student to solve this field, we
adopt a more practical approach and want to bring existing techniques to users.
%
In doing so, we seek to help users understand the solutions that program
synthesizers come up with.
%
Since current specifications tend to be underpowered, many unhelpful programs
come up from synthesizers.
%

Specifically, we seek to understand the degree to which input-output examples,
as properties of synthesized functions (instead of as specifications), aid users
in finding the program they came to the synthesizer for.

We base our work on \hoogleplus, a program synthesis engine for the
Haskell programming language \cite{hoogle_plus_2020}.
%
A user inputs a type signature of a function, like \texttt{a -> [Maybe a] -> a}
(read: ``A function that takes a thing and a list of optional-things, and
returns a thing'').
%
\hoogleplus then looks for Haskell programs that use library calls (from
a small, builtin set of functions) that satisfies the type signature.
%
Note that there may be an infinite number of candidate solutions that satisfy a type.
%
This presents a search problem for users of the system.

### 2020-03-20 methods.tex
grammar fixes

\paragraph{No-examples variant}
%
In the \noexamples variant, a user can enter a Haskell type signature at the top and get a
list of 10 candidate programs back.
%
These candidates are unmodified and in the same order that \hoogleplus
synthesizer generates them.
%
The results are deterministic meaning that for a given type signature, the
response is always the same.

\paragraph{Examples variant}
The \examples variant extends the \noexamples variant with input/output
examples for each candidate.
%
For each, type signature in the study, we manually created three simple
inputs of the appropriate type.
%
Each candidate program received the same three inputs.

\subsection{Procedure}
We gathered 10 haskell-literate participants from around our department's
building and from the local Hillcrest community.
%
Participants were each assigned to either the \noexamples treatment or the
\examples treatment for the duration of the study.
%
Each participant was presented a laptop with their respective tool variant open
and an interactive haskell interpreter (GHCI) open to the side.
%
Throughout the study participants were allowed to use only the tool and the
interpreter.

After a brief tutorial tasks, participants were asked to complete 4 tasks.
%
Each task consisted of a type signature they would enter, and an informal
english description of the behavior they were looking for.
%
They would then enter the type signature and identify the candidate program that
met the specification, if it was present.
%
Participants were told at the beginning that the solution may not be present
in the list---Task 3 did not have a solution in the list.

After completing the four tasks, the candidates were given a brief survey
that asked them about their haskell experience, their opinions on the tool,
and their strategies for finding the correct answer.

>>>>

\paragraph{No-examples variant}
%
In the \noexamples variant, a user can enter a Haskell type signature at the top and get a
list of 10 candidate programs back.
%
These candidates are mostly unmodified and in the same order that \hoogleplus
generates them.
%
The sole exception is Task 2, where the correct answer was manually inserted at
the end of the existing candidates.
%
The results are deterministic meaning that for a given type signature, the
response is always the same.

\paragraph{Examples variant}
The \examples variant extends the \noexamples variant with input/output
examples for each candidate.
%
For each type signature in the study, we manually created three simple
inputs of the appropriate type.
%
Each candidate program received the same three inputs.

\subsection{Procedure}
We gathered 10 Haskell-literate participants from around our department's
building and from the local Hillcrest community.
%
Participants were each assigned to either the \noexamples treatment or the
\examples treatment for the duration of the study.
%
Each participant was presented a laptop with their respective tool variant open
and an interactive Haskell interpreter (GHCI) open to the side.
%
Throughout the study participants were allowed to use only the tool and the
interpreter.

After a brief tutorial task, participants were asked to complete 4 tasks.
%
Each task consisted of a type signature and an informal
english description of the behavior they were looking for.
%
They would then enter the type signature and identify the candidate program that
met the specification, if it was present.
%
Participants were told at the beginning that the solution may not be present
in the list---Task 3 did not have a solution in the list.

After completing the four tasks, the candidates were given a brief survey
that asked them about their Haskell experience, their opinions on the tool,
and their strategies for finding the correct answer.

### 2020-03-20 results.tex
We have a methods section
grammar fixes

\section{Results}
\subsection{Time-to-correct solution}
Cumulatively, across the four tasks, we did not observe a significant change
in the time-to-correct solution.
%
A Wilcoxon Rank-Sum test, performed on the aggregate of per-task rankings,
yielded a confidence interval $p=0.14$.
%
Performing rank-sum of ranks is an especially conservative test, because it weighs
per-task rankings equally, regardless of the absolute times.

@@

\subsection{Incorrect selections}
When participants made an incorrect selection, they were allowed to keep
selecting until they found the correct program.
%
This happened 3 times in the \noexamples group and 5 times in the example group.
%
In the example group, a single participant made 3 incorrect guesses on a
single task.
%
This participant was also the only one who ranked their Haskell knowledge as a 1 out of 5.
%
Futhermore, the incorrect selections contained incorrect input/output pairs. 
%
With these anomalies in mind, we don't believe this study has determined a definitive answer
on the relationship between examples and error rates.

@@

% Figures.
\begin{figure*}[ht]
  \centering
  \includegraphics[width=\textwidth]{results/task_points.png}
  \caption{
    Examples did not noticeably affect time-to-completion, except in Task 2
    where the effect is dramatic.
  }
  \label{fig:data-points}
\end{figure*}

>>>>

\section{Results}
\begin{figure*}[ht]
  \centering
  \includegraphics[width=\textwidth]{results/task_points.png}
  \caption{
    Examples did not noticeably affect time-to-completion, except in Task 2
    where the effect is dramatic.
  }
  \label{fig:data-points}
\end{figure*}
\subsection{Time-to-correct solution}
Cumulatively, across the four tasks, we did not observe a significant change
in the time-to-correct solution.
%
A Wilcoxon Rank-Sum test, performed on the aggregate of per-task rankings,
yielded a confidence interval of $p=0.14$.
%
Performing rank-sum of ranks is an especially conservative test, because it weighs
per-task rankings equally, regardless of the absolute times or relative
difficulty of tasks.

@@

\subsection{Incorrect selections}
When participants made an incorrect selection, they were allowed to keep
selecting until they found the correct program.
%
This happened 3 times in the \noexamples group and 5 times in the example group.
%
In the example group, a single participant made 3 incorrect guesses on a
single task.
%
This participant was also the only one who ranked their Haskell knowledge as a 1 out of 5.
%
Furthermore, the incorrect selections contained incorrect input/output pairs.
%
With these anomalies in mind, we don't believe this study has determined a definitive answer
on the relationship between examples and error rates.

@@

% Figures.
